---
title: "Our World In Data: Life Expectancy"
format: html
editor: visual
---
## Import Data

```{r}
library(tidyverse)
library(mongolite)
library(janitor)
library(readxl)

rm(list=ls())
complete <- read_xlsx('marketing.xlsx',sheet='complete_cleaned') |> janitor::clean_names()
complete$life_expectancy10 <- as.numeric(complete$life_expectancy10)
complete$life_expectancy25 <- as.numeric(complete$life_expectancy25)
complete$life_expectancy45 <- as.numeric(complete$life_expectancy45)
complete$life_expectancy65 <- as.numeric(complete$life_expectancy65)
complete$life_expectancy80 <- as.numeric(complete$life_expectancy80)
complete$life_expectancy_diff_fm <- as.numeric(complete$life_expectancy_diff_fm)
complete$ymd <- as.Date(complete$ymd)

str(complete)
```

## Requrements:
### List of Country Only Data

```{r}
 countries <- complete %>% clean_names() |> filter(!code %in% c("NA","OWID_"))
str(countries)
```

### Average Life Expectancy and Male and Female Differences by Year

```{r}
countries_life_expectancy <- countries |> group_by(ymd) |>
        summarize( average_le = mean(life_expectancy),
        average_female_male = mean(life_expectancy_diff_fm))
        head(countries_life_expectancy)

countries_life_expectancy |> ggplot(aes(x=ymd,y=average_le)) + geom_line() + labs(title ="Yearly Average Life expectancy")
countries_life_expectancy |> ggplot(aes(x=ymd,y=average_female_male)) + geom_line() + labs(title ="Yearly Average Difference Male/Female")
```
### Create marketing MongoDB Db and three table  from result 
```{r}
conn <- mongo(collection = "countries", db = "marketing", url = "mongodb://localhost")
conn$drop() 
conn$insert(countries)
results <- conn$find('{}') 
str(results)
```
```{r}
conn1 <- mongo(collection = "countries_life_expectancy", db = "marketing", url = "mongodb://localhost")
conn1$drop() 
conn1$insert(countries_life_expectancy)
results <- conn1$find('{}') 
str(results)
```

## Countries Listed as Low-income countries, Middle-income countries and High-income countries

```{r}
 income_groups <- complete  |> filter(entity %in% c("Low-income countries","Middle-income countries","High-income countries")) |> 
        select(entity,ymd,life_expectancy,life_expectancy_diff_fm)
str(income_groups)

ggplot(income_groups) +geom_line(aes(x=ymd,y=life_expectancy,col=entity))
ggplot(income_groups) +geom_line(aes(x=ymd,y=life_expectancy_diff_fm,col=entity))

```
```{r}
conn2 <- mongo(collection = "income_groups", db = "marketing", url = "mongodb://localhost")
conn2$drop() 
conn2$insert(income_groups)
results <- conn2$find('{}') 
str(results)
```
## Average for life_expectancy and life_expectancy_diff_fm

```{r}

income_group_by_year <- income_groups |> group_by(ymd) |> summarize( average_le = mean(life_expectancy),
        average_female_male = mean(life_expectancy_diff_fm))
        head(income_group_by_year)

```
```{r}
ggplot(income_group_by_year) + geom_line(aes(x=ymd,y=average_le))
ggplot(income_group_by_year) + geom_line(aes(x=ymd,y=average_female_male))

```
```{r}
conn3 <- mongo(collection = "income_group_by_year", db = "marketing", url = "mongodb://localhost")
conn3$drop() 
conn3$insert(income_group_by_year)
results <- conn1$find('{}') 
str(results)
```
## First and Last Life Expectancy
```{r}

# Assuming 'countries' is your data frame and it has columns 'entity' for country and 'life_expectancy'
countries_by_change_in_life_expectancy <- countries %>%
  group_by(entity) %>%
  arrange(entity, ymd) %>%  # Assuming you want to calculate the change over time
  summarize(
    first_life_expectancy = round(first(life_expectancy),2),
    last_life_expectancy = round(last(life_expectancy),2),
    percent_change = round(((last(life_expectancy) - first(life_expectancy)) / first(life_expectancy)) * 100, 2)
  )
```

```{r}
library(tidyverse)

# Assuming 'countries' is your data frame and it has columns 'entity', 'ymd', and 'life_expectancy'
life_expectancy_data <- countries %>%
  group_by(entity) %>%
  arrange(entity, ymd) %>%
  summarize(
    first_year = first(ymd),
    last_year = last(ymd),
    first_life_expectancy = first(life_expectancy),
    last_life_expectancy = last(life_expectancy)
  ) %>%
  pivot_longer(
    cols = c(first_life_expectancy, last_life_expectancy),
    names_to = "time_point",
    values_to = "life_expectancy"
  ) %>%
  mutate(time_point = ifelse(time_point == "first_life_expectancy", "First", "Last")) |> head(6)

# Plotting with wider lines
ggplot(life_expectancy_data, aes(x = time_point, y = life_expectancy, group = entity)) +
  geom_line(size = 3) +  # Increase the size for wider lines
  geom_point() +
  facet_wrap(~ entity, scales = "free_y") +
  theme_minimal() +
  labs(title = "First and Last Life Expectancy Values for Each Country",
       x = "Time Point",
       y = "Life Expectancy")


```{r}
# conn4 <- mongo(collection = "countries_by_change_in_life_expectancy", db = "marketing", url = "mongodb://localhost")
# conn4$drop() 
# conn4$insert(countries_by_change_in_life_expectancy)
# countries_by_change_in_life_expectancy <- conn1$find('{}') 
# countries_by_change_in_life_expectancy$ymd <- as.Date(countries_by_change_in_life_expectancy$ymd)
# str(countries_by_change_in_life_expectancy)
```
